---
title: "Stage 5.2 — Zonation inputs"
output: html_document
params:
  config_file: "config/madagascar.yml"
  mode: "inspect"              # inspect | inputs | feature_list
---

## 1. Configuration and preflight

```{r setup, message=TRUE, warning=FALSE}
source(file.path("R", "stage52_workflow.R"))
rmd_options <- project_rmd_params("5.2_build_binary_patch_rasters_zonation_feature_list.Rmd")
configured <- rmd_project_config(rmd_options, "stage52")
rmd_params <- configured$params
context <- application_rmd_context(rmd_params)
config <- stage52_application_config(rmd_params, context)
describe_stage52_config(config)
```

## 2. Validate inputs

`inputs` uses the current patch lookup and patch rasters to rebuild the
complete binary feature set. `feature_list` treats the top-level TIFFs already
in `Patches_binary` as authoritative and refreshes only the feature list and
its derived Zonation settings file. `inspect` reports status without reading
raster values.

## 3. Run selected operation

```{r build_zonation_inputs, message=TRUE}
if (identical(config$mode, "inspect")) {
  stage52_status <- inspect_stage52(config)
  print(stage52_status)
} else {
  stage52_result <- run_stage52(config)
  # Only the full input build is a durable pipeline boundary. The lightweight
  # path refresh must not alter the application recovery state.
  if (identical(config$mode, "inputs")) {
    application_commit_boundary(
      context, application_stage52_boundary(config, stage52_result)
    )
  }
}
```

## 4. Commit outputs

`inputs` commits the feature list and settings file with its rebuilt binary
rasters. `feature_list` atomically replaces only those two text files.
Completed Zonation rank maps remain in their method output directories and are
read there by Stage 7.2.

## 5. Summary

```{r summary, message=TRUE}
if (identical(config$mode, "inspect")) {
  log_msg(
    "Stage 5.2 | inspection complete | available=",
    sum(stage52_status$exists), "/", nrow(stage52_status),
    "| files_written=none"
  )
} else if (identical(config$mode, "inputs")) {
  log_msg(
    "Stage 5.2 | outputs | binary_rasters=", length(stage52_result$binary_paths),
    "| method_dirs=", length(stage52_result$method_dirs),
    "| feature_list=", stage52_result$feature_list,
    "| settings=", stage52_result$settings_file
  )
} else {
  log_msg(
    "Stage 5.2 | outputs | features=", stage52_result$feature_count,
    "| binary_rasters_written=0",
    "| feature_list=", stage52_result$feature_list,
    "| settings=", stage52_result$settings_file
  )
}
```
