---
title: "Stage 5.3 — Initial persistence"
output: html_document
params:
  config_file: "config/madagascar.yml"
  curve: "q50"                 # q025 | q16 | q50 | q84 | q975
---

```{r setup, message=TRUE, warning=FALSE}
source(file.path("R", "stage53_workflow.R"))
rmd_options <- project_rmd_params("5.3_initial_species_persistence.Rmd")
configured <- rmd_project_config(rmd_options, "stage53")
rmd_params <- configured$params
context <- application_rmd_context(rmd_params)
config <- stage53_application_config(rmd_params, context)
describe_stage53_config(config)

initial_report <- run_stage53(config)
application_commit_boundary(
  context, application_stage53_boundary(config, initial_report)
)
application_finalize_scenario(context, require_complete = FALSE)
```

## Interpretation

This report describes the selected taxa in the final Stage 5 landscape before
Stage 6 removes any cells. Population-unit persistence uses the selected
shifted-Gompertz curve. The tables are displayed during interactive chunk
execution and knitting; the report also writes the corresponding SI figure.
Species persistence is `1 - product(1 - P(PU))`, treating population-unit
extinction outcomes as independent. Species without a retained Stage 5
population unit are coverage gaps and are not assigned persistence zero.

## Analysis coverage

```{r coverage, echo=FALSE}
print(knitr::kable(initial_report$coverage, digits = c(NA, 0, 0, 0, 1),
                   caption = "Selected and Stage 5-represented species."))
if (nrow(initial_report$absent_species)) {
  print(knitr::kable(
    utils::head(initial_report$absent_species, 10L),
    caption = "Selected species without retained Stage 5 population units."
  ))
}
```

## Population-unit summary

```{r pu_summary, echo=FALSE}
print(knitr::kable(initial_report$pu_summary, digits = 3,
                   caption = "Initial population-unit area and persistence statistics."))
```

## Species summary

```{r species_summary, echo=FALSE}
print(knitr::kable(initial_report$species_summary, digits = 3,
                   caption = "Initial species-level persistence statistics."))
```

## Area-persistence association

Spearman correlations are descriptive because species differ in density,
demographic parameters, and the distribution of area among population units.

```{r correlations, echo=FALSE}
print(knitr::kable(initial_report$correlations, digits = 4,
                   caption = "Spearman association between retained PU area and species persistence."))
```

## Species extremes

```{r species_extremes, echo=FALSE}
print(knitr::kable(initial_report$ranked_species$lowest, digits = 4,
                   caption = "Lowest-persistence species (up to 10)."))
print(knitr::kable(initial_report$ranked_species$highest, digits = 4,
                   caption = "Highest-persistence species (up to 10)."))
```

## Initial persistence figure

In panel B, point color identifies taxon and point shape identifies the SDM
branch used for each species (PPM or RangeBag).

```{r initial_persistence_figure, echo=FALSE, fig.width=7.0, fig.height=3.4, dpi=300}
print(initial_report$figure)
```

```{r summary, message=TRUE}
log_msg(
  "Stage 5.3 | curve=", initial_report$curve,
  "| taxa=", initial_report$taxa,
  "| represented_species=", nrow(initial_report$species),
  "| population_units=", nrow(initial_report$pu),
  "| median_species_persistence=",
  signif(stats::median(initial_report$species$species_persistence), 4),
  "| figure=", config$paths$figure
)
```
