---
title: "Stage 5 — Patches and connectivity"
output: html_document
params:
  config_file: "config/madagascar.yml"
  mode: "inspect"              # inspect | resume | restart
  verbose: false
  clump_backend: "auto"        # auto | fasterRaster | terra
  grass_dir: null
---

## 1. Configuration and preflight

```{r setup, message=TRUE, warning=FALSE}
source(file.path("R", "stage5_workflow.R"))
rmd_options <- project_rmd_params("5_build_patches_and_connectivity.Rmd")
configured <- rmd_project_config(rmd_options, "stage5")
rmd_params <- configured$params
context <- application_rmd_context(rmd_params)
landcover_fingerprint <- if (!identical(rmd_params$mode, "inspect")) {
  validate_application_rmd_spatial_inputs(context, rmd_params)
} else NULL
config <- stage5_application_config(
  rmd_params, context, landcover_fingerprint = landcover_fingerprint
)
describe_stage5_config(config)
```

## 2. Read and select species

The Stage 4 table is validated and filtered by `taxa`. This selection does not
inspect or require Gompertz coefficient columns.

## 3. Inspect recovery state or initialize spatial work

`inspect` reports input, final-artifact, and checkpoint presence without
loading spatial packages, deserializing large graphs, hashing SDMs, resolving
the clumping backend, or reading land-cover cell values. Active modes perform
exact compatibility validation before creating the ROI-cropped habitat masks,
common template, and reusable cell-area raster.

## 4. Process or reuse species

```{r run_pipeline, message=TRUE}
stage5_result <- run_stage5(config)
if (!identical(config$mode, "inspect")) {
  application_commit_boundary(
    context, application_stage5_boundary(config, stage5_result)
  )
}
```

## 5. Validate and commit outputs

Each successful scientific outcome is validated and committed independently to
the recovery directory. Final rasters, lookup, connectivity, and provenance
metadata replace the prior output set only after every selected species has a
valid checkpoint and the assembled objects pass cross-object validation. A new
checkpoint's validated result is reused immediately; a checkpoint loaded in a
later session still receives full raster, lookup, and connectivity validation.

## 6. Final diagnostics

```{r summary, message=TRUE}
report_stage5(stage5_result)
if (!is.null(stage5_result$runtime_log_path)) {
  cat("Runtime log:", stage5_result$runtime_log_path, "\n")
}
```
