---
title: "Stage 7.1 — Priority results"
output: html_document
params:
  config_file: "config/madagascar.yml"
  curve: "q50"                # q025 | q16 | q50 | q84 | q975
---

```{r setup, message=FALSE, warning=FALSE}
# Stage 7.1 is a deterministic reporting stage. It identifies one completed
# Stage 6 run, reconstructs its ecological states, validates its removal
# surface, and atomically regenerates the associated table and figures.
source(file.path("R", "stage71_workflow.R"))

rmd_options <- project_rmd_params("7.1_stage_meta_and_priority_surface.Rmd")
configured <- rmd_project_config(rmd_options, "stage71")
rmd_params <- configured$params
context <- application_rmd_context(
  rmd_params, require_application = TRUE, require_scenario = TRUE
)
config <- stage71_application_config(rmd_params, context)
log_stage7_config(config)
```

## Reconstruct and validate the priority results

Stage zero is read from the shared Stage 6 initialization. Every positive stage is read from
its corresponding patch lookup. Persistence is not interpolated between
completed stages. The same spatial states are also evaluated under all five
demographic coefficient sets.

```{r calculate}
result <- run_stage71(config)
application_commit_boundary(
  context, application_stage71_boundary(config, result)
)
report_tables <- stage71_report_tables(result)
```

## Assemblage summaries

```{r persistence-overview, echo=FALSE}
assert(
  exists("result"),
  "Run the Stage 7.1 calculation chunk before displaying its report."
)

state_summary <- report_tables$state_summary
print(knitr::kable(state_summary, digits = 4, row.names = FALSE))
```

## Species summaries

### Largest initial-to-latest persistence losses

```{r largest-losses, echo=FALSE}
largest_losses <- report_tables$largest_losses
print(knitr::kable(largest_losses, digits = 4, row.names = FALSE))
```

### Smallest area under the persistence trajectory

The raw AUC integrates persistence against the percentage of initially alive
cells removed. The normalized AUC divides by the observed removal span and
lies in `[0,1]`.

```{r smallest-auc, echo=FALSE}
smallest_auc <- report_tables$smallest_auc
print(knitr::kable(smallest_auc, digits = 4, row.names = FALSE))
```

### Lowest persistence at the latest completed stage

```{r lowest-latest, echo=FALSE}
lowest_latest <- report_tables$lowest_latest
print(knitr::kable(lowest_latest, digits = 4, row.names = FALSE))
```

### Species with no surviving population unit

```{r species-without-pus, echo=FALSE}
without_pus <- report_tables$without_pus
if (nrow(without_pus)) {
  print(knitr::kable(without_pus, digits = 4, row.names = FALSE))
} else {
  cat("none\n")
}
```

## Supplementary figures

### Assemblage persistence

```{r community-trajectory, echo=FALSE, out.width='100%'}
print(knitr::include_graphics(
  result$outputs[["pipeline_persistence"]]
))
```

### Demographic-curve uncertainty

```{r mean-curve-uncertainty, echo=FALSE, out.width='100%'}
print(knitr::include_graphics(
  result$outputs[["curve_uncertainty"]]
))
```

### Species persistence

```{r species-heatmap, echo=FALSE, out.width='100%'}
print(knitr::include_graphics(
  result$outputs[["species_persistence"]]
))
```

## Continuous cell-removal order

Lower values identify cells removed earlier; higher values identify cells
retained longer. The map is built only after the integer removal raster,
continuous rank map, event table, and initial Stage 6 domain have been shown
to agree.

```{r removal-order-map, echo=FALSE, out.width='70%'}
print(knitr::include_graphics(
  result$outputs[["cell_removal_order"]]
))
```

## Final summary

```{r summary, echo=FALSE}
cat("Run ID: ", config$run_id, "\n", sep = "")
cat("Primary persistence curve: ", config$curve, "\n", sep = "")
cat("Ecological states: ", nrow(result$stage_meta), "\n", sep = "")
cat(
  "Retained species: ",
  data.table::uniqueN(result$persistence$species$species),
  "\n",
  sep = ""
)
cat(
  "Ranked cells: ",
  format(result$surface$n_ranked, big.mark = ","),
  "\n",
  sep = ""
)
cat("Outputs:\n")
cat(paste0("  - ", unname(result$outputs)), sep = "\n")
cat("\nElapsed seconds: ", sprintf("%.1f", result$elapsed), "\n", sep = "")
```
