---
title: "Stage 7.3 — Persistence comparison"
output: html_document

params:
  config_file: "config/madagascar.yml"
  mode: "inspect"             # inspect | report
  main_curve: "q50"
  rank_method: "abf"          # abf | caz1 | caz2 | cazmax
---

```{r setup, message=FALSE, warning=FALSE}
# Stage 7.3 compares one completed Stage 6 trajectory with the corresponding
# Stage 7.2 benchmark lookup sequence. Paths are derived from the selected
# application, threshold, removal schedule, curve, and benchmark.
source(file.path("R", "stage73_workflow.R"))

rmd_options <- project_rmd_params("7.3_persist_cmp.Rmd")
configured <- rmd_project_config(rmd_options, "stage73")
rmd_params <- configured$params
config <- stage73_application_config(rmd_params)
```

## Configuration and artifact status

`inspect` reports paths without opening scientific objects or writing files.
Active modes regenerate the inexpensive comparison from pipeline states and
the exact retained-cell benchmark lookup library.

```{r inspect}
if (identical(config$mode, "inspect")) {
  inspection <- inspect_stage73(config)
  print(inspection)
}
```

## Calculate the comparison

The persistence-based and benchmark methods use corresponding completed stage
numbers. All retained species remain in the denominator at every stage;
species absent from a spatial state receive persistence zero.

```{r calculate}
if (!identical(config$mode, "inspect")) {
  result <- run_stage73(config)
}
```

## Seven-panel persistence comparison

Panel `(a)` summarizes the assemblage over the complete removal sequence.
Panels `(b)`–`(g)` show persistence, retained population-unit composition, and
the persistence–area relationship for the two configured focal species.

```{r combined-figure, echo=FALSE, out.width='100%'}
if (!identical(config$mode, "inspect")) {
  assert(exists("result"), "Run the calculation chunk before displaying the figure.")
  print(knitr::include_graphics(result$figure_path))
}
```

## Key comparison metrics

### Species below the persistence threshold at late-stage checkpoints

Classification uses strict
\(S < `r config$persistence_threshold`\). Each target uses the completed
ecological stage nearest 90%, 95%, or 99% removal.

```{r metrics-late-stage, echo=FALSE}
if (!identical(config$mode, "inspect")) {
  print(knitr::kable(
    result$report_tables$late_stage_threshold,
    digits = c(1, 1, 0, 0, 0, 0, 0, 0),
    align = "rrrrrrrr"
  ))
}
```

### Species favoring each method

The removal-sequence comparison integrates the persistence difference over
the complete cell-removal axis. The retained-area comparison integrates over
each species' shared retained-PU-area support.

```{r metrics-favoring, echo=FALSE}
if (!identical(config$mode, "inspect")) {
  print(knitr::kable(
    result$report_tables$favoring_by_scope,
    digits = c(NA, NA, 0, 0, 1, 0, 1, 0, 1, 1),
    align = "llrrrrrrrr"
  ))
}
```

### Three largest positive sequence-averaged differences

```{r metrics-top3, echo=FALSE}
if (!identical(config$mode, "inspect")) {
  print(knitr::kable(
    result$report_tables$top3_positive,
    digits = c(0, NA, NA, 3, 3, 1),
    align = "rllrrr"
  ))
}
```

### Sensitivity across the five persistence functions

The same spatial states are evaluated with q2.5, q16, q50, q84, and q97.5
coefficient pairs. The first table summarizes checkpoint directions; the
second summarizes species-level direction stability across the five curves.

```{r metrics-sensitivity-checkpoints, echo=FALSE}
if (!identical(config$mode, "inspect") &&
    nrow(result$report_tables$sensitivity_checkpoints)) {
  print(knitr::kable(
    result$report_tables$sensitivity_checkpoints,
    digits = c(1, 1, 3, 0, 0, NA),
    align = "rrrrrl"
  ))
}
```

```{r metrics-sensitivity-stability, echo=FALSE}
if (!identical(config$mode, "inspect") &&
    nrow(result$report_tables$sensitivity_stability)) {
  print(knitr::kable(
    result$report_tables$sensitivity_stability,
    digits = c(NA, 0),
    align = "lr"
  ))
}
```

## Output summary

```{r summary, echo=FALSE}
if (!identical(config$mode, "inspect")) {
  cat("Run ID: ", config$run_id, "\n", sep = "")
  cat("Benchmark: ", config$rank_label, "\n", sep = "")
  cat("Benchmark lookup directory: ", config$paths$lookup_library$lookups, "\n", sep = "")
  cat("Persistence threshold: ", config$persistence_threshold, "\n", sep = "")
  cat("Focal species: ", paste(config$focal_species, collapse = "; "), "\n", sep = "")
  cat("Combined figure: ", result$figure_path, "\n", sep = "")
}
```
