---
title: "Stage 8 — Selective pipeline orchestration"
output: html_document
params:
  config_file: "config/madagascar.yml"
  # Run control. inspect never creates or changes analysis files.
  mode: "resume"                   # inspect | resume | restart
  # Core scientific work only. Add 5.1, 5.2, 5.3, or 7.1 explicitly when
  # those explanatory, external-handoff, or reporting products are wanted.
  stages:
    value:
      - "stage_2"
      - "stage_3"
      - "stage_4"
      - "stage_5"
      - "stage_5_1"
      - "stage_5_3"
      - "stage_6"
      - "stage_7_1"
  restart_from: null                 # null | stage_2 | stage_4 | stage_6
  stage3_figures: true               # applies only when stage_3 is selected
  process_figure_species: "Cryptoprocta ferox"
                                     # required only when Stage 5.1 is selected

  # Stage 5 clumping implementation. auto prefers fasterRaster/GRASS.
  clump_backend: "fasterRaster"      # auto | fasterRaster | terra
  grass_dir: null                    # null = environment/standard discovery

  # Stage 6 runs once for each requested curve. q50 is shown when available.
  priority_curves:
    value: ["q025", "q16", "q50", "q84", "q975"]

  threads: null

  max_stages: .inf
  checkpoint_every_stages: 1
---

## How to use this driver

Use `inspect` first to see selected work, prerequisite handoffs, and completed
artifacts. `stages` is literal: an omitted prerequisite is validated and reused,
never executed automatically. Use `resume` for new or compatible work. With
`restart`, set `restart_from` to one of the deliberately coarse, transactionally
safe boundaries. A lean calculation without optional products can select
Stages 2, 3, 4, 5, and 6 and set `stage3_figures: false`.

## Configuration

```{r configuration, message=TRUE, warning=FALSE}
source(file.path("R", "stage8_workflow.R"))

rmd_options <- project_rmd_params("8_run_pipeline.Rmd")
configured <- rmd_project_config(rmd_options, "stage8")
rmd_params <- configured$params
config <- stage8_config(rmd_params)
cat("Persistence model:", config$model$root, "\n")
cat("Application scenario:", config$scenario$root, "\n")
cat("Priority run:", config$scenario$run_root, "\n")
print(knitr::kable(stage8_settings_table(config), row.names = FALSE))
input_inspection <- inspect_project_config(configured$configuration)
print(knitr::kable(input_inspection$inputs, row.names = FALSE))
print(knitr::kable(input_inspection$sdm_directories, row.names = FALSE))
```

## Run or inspect

`inspect` is read-only. `resume` validates completed boundaries and continues
selected work. `restart` replaces only `restart_from`: the reusable persistence
model and current scenario at `stage_2`, the current threshold scenario at
`stage_4`, or the requested curve directories at `stage_6`.

```{r run, message=TRUE, warning=FALSE}
stage8_result <- run_stage8(config)
```

## Execution plan

```{r plan, echo=FALSE}
print(knitr::kable(stage8_result$plan, row.names = FALSE))
```

## Stage status

```{r status, echo=FALSE}
print(knitr::kable(stage8_result$status, row.names = FALSE))
if (!isTRUE(stage8_result$complete) && !is.null(stage8_result$stop_reason)) {
  cat("\n", stage8_result$stop_reason, "\n", sep = "")
}
```

## Artifacts

```{r artifacts, echo=FALSE}
if (nrow(stage8_result$artifacts)) {
  shown <- summarize_application_artifacts(stage8_result$artifacts)
  shown$total_bytes <- format(shown$total_bytes, big.mark = ",", scientific = FALSE)
  print(knitr::kable(shown, row.names = FALSE))
  cat("\nThe complete file-level table is available as `stage8_result$artifacts`.\n")
} else {
  cat("No completed artifacts are recorded for this analysis.\n")
}
```

## Initial persistence summary

```{r stage53-summary, echo=FALSE}
if (!is.null(stage8_result$stage53)) {
  print(knitr::kable(stage8_result$stage53$coverage, digits = 2, row.names = FALSE))
  print(knitr::kable(stage8_result$stage53$species_summary, digits = 3, row.names = FALSE))
}
```

## Priority-result summary

```{r stage71-summary, echo=FALSE}
if (!is.null(stage8_result$stage71)) {
  print(knitr::kable(stage8_result$stage71$initial, digits = 4, row.names = FALSE,
                     caption = "Initial Stage 7.1 state."))
  print(knitr::kable(stage8_result$stage71$latest, digits = 4, row.names = FALSE,
                     caption = "Latest Stage 7.1 state."))
  print(knitr::kable(stage8_result$stage71$lowest_auc, digits = 4, row.names = FALSE,
                     caption = "Species with the lowest normalized persistence AUC."))
}
```

## Generated figures

```{r figures, echo=FALSE, fig.width=8, fig.height=6}
figure_paths <- stage8_result$figures
figure_paths <- figure_paths[file.exists(figure_paths)]
if (length(figure_paths)) {
  for (path in figure_paths) {
    print(knitr::include_graphics(path))
  }
} else {
  cat("No selected Stage 8 figures are available yet.\n")
}
```
